---
title: "R Notebook"
output:
  html_notebook: default
  html_document:
    toc: yes
    toc_float: yes
---

# Intro

Starting to dive into the data!

# Setup

## Load Packages
```{r}
library(ggplot2)
library(dplyr)
library(tm)
library(stringr)
library(jsonlite)
library(googlesheets)
library(randomForest)
```


## Load data
```{r import clean csv, cache = T, results = "hide"}
raw_data <- read.csv("./data/CLEAN-DATA.csv", stringsAsFactors = F)
CleanDataKeyEntry <- gs_title("CleanDataKeyEntry")
main_data <- gs_read(CleanDataKeyEntry, ws = "MainData")
```

# NTEE Cleaning
In a quick peek of the data, we see the NTEE code column has `r round(sum(raw_data$NTEE_Code=="")/nrow(raw_data)*100,2)` percent missing. Multiple codes are allowed, so we will set out on imputing some missing values.

## Parsing JSON NTEE codes to data frame
In order to join the NTEE code list to the data, we need to parse the JSON file to a data frame. And bonus: add it to the Google sheet! Note that not all codes have keywords assocaited with them. Keywords are separated with "|". 
```{r NTEE json to df, cache = T}
ntee <- fromJSON("./data/ntee.json", simplifyDataFrame = T, flatten = T)
ntee_df <- data.frame(code = c(),
                    title = c(),
                    desc = c(),
                    keywords = c()
                    )
for (i in 1:length(ntee)) {
  code_add      <- names(ntee)[i]
  code_data     <- ntee[[i]]
  code_desc     <- code_data$description
  code_keywords <- paste(code_data$keywords, collapse = "|")
  code_title    <- code_data$title
  
  data_add <- data.frame(code     = code_add,
                         title    = code_title,
                         desc     = code_desc,
                         keywords = code_keywords
                         )
  ntee_df <- rbind(ntee_df, data_add)
}

# CleanDataKeyEntry <- CleanDataKeyEntry %>%
#   gs_ws_new(ws_title = "NTEE Codes", input = ntee_df)
```

## Create table of NPO name and and codes
The aid with the cleaning, we'll reduce the raw data to a table of NPO names and associated codes.

*work paused on this*


# NTEE Top Level Feature Creation
Here we parse the assigned NTEE codes and extract the top level (i.e. A, B, C) code and one-hot encode the results. 

# Give Reason Analysis
Let's see if we can do a basic model of giving reason.
Giving reasons: 1 = Personal, 2 = Invested, 3 = Exposure, 4 =  ??, NA

## Data Cleaing and Subsetting
Since not all NPOs matched well to the Guidestar data, we'll follow the convention of filtering out observations that were not good matches using *Inclusion == TRUE*

```{r gr data prep, cache = T}
gr_data <- main_data %>%
  filter(Inclusion == TRUE)

# Recode give reason
## Recoding gr_data$give_reason into gr_data$give_reason2
## Recoding gr_data$give_reason_vlook into gr_data$give_reason2
gr_data$give_reason2 <- gr_data$give_reason_vlook
gr_data$give_reason2[gr_data$give_reason_vlook == "1"] <- "Personal"
gr_data$give_reason2[gr_data$give_reason_vlook == "2"] <- "Invested"
gr_data$give_reason2[gr_data$give_reason_vlook == "3"] <- "Exposure"
gr_data$give_reason2[gr_data$give_reason_vlook == "#N/A"] <- "Not Found"
gr_data$give_reason2[gr_data$give_reason_vlook == "??"] <- "Not Found"
gr_data$give_reason2[gr_data$give_reason_vlook == "4"] <- "Other"
gr_data$give_reason2[is.na(gr_data$give_reason_vlook)] <- "Not Found"
gr_data$give_reason2 <- factor(gr_data$give_reason2)


```

## Frequency plot 
A quick check of frequency of give reason.
Filtered for the 3 main give reasons.
```{r gr freq plot}
gr_data %>%
  filter(give_reason_vlook %in% c("1", "2", "3")) %>%
  group_by(give_reason2) %>%
  summarise(freq = n()) %>%
  ggplot(aes(give_reason2, freq)) +
    geom_bar(stat = "Identity")

gr_data %>%
  filter(give_reason_vlook %in% c("1", "2", "3")) %>%
  group_by(give_reason2) %>%
  summarise(avg_local_count = mean(votes_local_network, na.rm = T)) %>%
  ggplot(aes(give_reason2, avg_local_count)) +
  geom_bar(stat = "Identity") +
  ggtitle("Average Local Network Votes by Give Reason")
```
## Give Reason Freq by Photo
Plot of give reason frequncy by if Photo was included or not.
```{r gr photo freq}
gr_data %>%
  group_by(give_reason2, Photo) %>%
    summarise(freq = n()) %>%
  ggplot(aes(give_reason2, freq)) +
    geom_bar(stat = "Identity") +
  facet_wrap(~ Photo)
```

## Give Reason Freq by top level email domain
```{r gr tl domain freq}
gr_data %>%
  group_by(give_reason2, EmailTLD) %>%
    summarise(freq = n()) %>%
  ggplot(aes(give_reason2, freq)) +
    geom_bar(stat = "Identity") +
  facet_wrap(~ EmailTLD) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```
A bit sparse for somedomains, so let's recode as *com = 1/0* where 1 = "was .com" and 0 = "not .com".
```{r emailLTD recode}
## Recoding gr_data$EmailTLD into gr_data$com
gr_data$com <- gr_data$EmailTLD
gr_data$com[gr_data$EmailTLD == "com"] <- "1"
gr_data$com[gr_data$EmailTLD == "net"] <- "0"
gr_data$com[gr_data$EmailTLD == "in"] <- "0"
gr_data$com[gr_data$EmailTLD == "edu"] <- "0"
gr_data$com[gr_data$EmailTLD == "org"] <- "0"
gr_data$com[gr_data$EmailTLD == "gov"] <- "0"
gr_data$com[gr_data$EmailTLD == "de"] <- "0"
gr_data$com[gr_data$EmailTLD == "us"] <- "0"
gr_data$com[gr_data$EmailTLD == "nl"] <- "0"
gr_data$com <- as.numeric(gr_data$com)

gr_data %>%
  group_by(give_reason2, com) %>%
    summarise(freq = n()) %>%
  ggplot(aes(give_reason2, freq)) +
    geom_bar(stat = "Identity") +
  facet_wrap(~ com) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))
```

## Give Reason freq by high level NTEE
```{r gr ntee freq}
# gr_data <- gr_data %>%
#   mutate()


```




##  Multinomal Model of Give Reason
A basic multinomial model if give reason were we consider give reason == NA as a valid choice (i.e. choice not to disclose either purposefully or not)

```{r gr multinom, cache= T}
library(nnet)

gr_model_1 <- gr_data %>% 
 multinom(give_reason2 ~ Photo + give_method + com, data = .)

summary(gr_model_1)
```



# Vote Count Prediction
Let's see if we can build a predictive model for vote count. What factors are predictive of story vote count. Since we expect there to be a bit of a network effect with people that have a large number of friends getting large local vote counts, we will use gallery_count/local_count as the response variable. To start, we'll simply focus on getting a feature impotance list.

## Feature Extraction and Model Building
Let's extract some key features.
```{r gallery vote ratio model}
# Extract columns to keep
set.seed(101)
keep_cols <- c("votes_gallery_(Calc)"
               ,"votes_local_network"
               ,"InferredGender"
               ,"TextLength"
               ,"NumWords"
               ,"NumPunctuation"
               ,"NumURLs"
               ,"NumHashtags"
               ,"NumAtMentions"
               ,"NumDollarsigns"
               ,"NumDigits"
               ,"NumExclamationpoints"
               ,"NumUppercaseletters"
               ,"EmailTLD"
               ,"ScrabbleScore"
               ,"Total Revenue"
               ,"Photo"
               ,"give_reason2"
               #,"give_method_vlookup"
               ,"give_method"
               ,"SentimentScore"
               ,"EmailDomain"
               )

vm_data <- gr_data %>%
  select(one_of(keep_cols))

# Target
vm_data$target <- round(vm_data$`votes_gallery_(Calc)`/vm_data$votes_local_network,2)

# Some Cleaning

vm_data$total_revenue <- as.numeric(gsub("\\,", '', sub('\\$','',as.character(vm_data$`Total Revenue`))))

vm_data$sentiment_clean <- round(as.numeric(vm_data$SentimentScore),5)

vm_data$InferredGender <- as.factor(vm_data$InferredGender)
vm_data$EmailTLD <- as.factor(vm_data$EmailTLD)
vm_data$Photo <- as.factor(vm_data$Photo)
vm_data$EmailDomain <- as.factor(vm_data$EmailDomain)
#vm_data$give_method_vlookup <- as.factor(vm_data$give_method_vlookup)
vm_data$give_method <- as.factor(vm_data$give_method)
# Drop old features

drop_cols <- c("SentimentScore"
               ,"Total Revenue"
               ,"votes_gallery_(Calc)"
               ,"votes_local_network")

vm_data <- vm_data %>% select(-one_of(drop_cols))

# Remove NA's in respose
vm_data <- vm_data %>% filter(!is.na(target))

vm_data2 <- na.roughfix(vm_data)

# Basic GBM for importance

vm_gbm <- gbm(log(target+1) ~.
              ,data = vm_data2
              ,distribution = "gaussian"
              ,n.trees = 1000
              ,interaction.depth = 5
              ,n.minobsinnode = 15
              ,shrinkage = 0.005
              ,bag.fraction = 0.5
              ,train.fraction = 0.7
              ,verbose = TRUE)

summary(vm_gbm)

gbm.perf(vm_gbm)


```


```{r email domain freq}
vm_data2 %>% group_by(EmailDomain) %>%
  summarise(freq = n()) %>%
  filter(freq >= 25) %>%
  ggplot(aes(EmailDomain, freq)) + 
  geom_bar(stat="identity")

vm_data2 %>% group_by(EmailDomain) %>%
  summarise(avg_target = mean(target)
            ,freq = n()) %>%
  filter(freq >= 25) %>%
  ggplot(aes(EmailDomain, avg_target)) + 
  geom_bar(stat="identity")

```



## Modelinh gallery count directly
Let's extract some key features.
```{r gallery vote count model}
set.seed(101)
# Extract columns to keep
keep_cols <- c("votes_gallery_(Calc)"
               ,"InferredGender"
               ,"TextLength"
               ,"NumWords"
               ,"NumPunctuation"
               ,"NumURLs"
               ,"NumHashtags"
               ,"NumAtMentions"
               ,"NumDollarsigns"
               ,"NumDigits"
               ,"NumExclamationpoints"
               ,"NumUppercaseletters"
               ,"EmailTLD"
               ,"ScrabbleScore"
               ,"Total Revenue"
               ,"Photo"
               ,"give_reason2"
               #,"give_method_vlookup"
               ,"give_method"
               ,"SentimentScore"
               ,"EmailDomain"
               )

vm_data <- gr_data %>%
  select(one_of(keep_cols))

# Target
vm_data$target <- vm_data$'votes_gallery_(Calc)'

# Some Cleaning
# Use total revenue on log scale due to some biiiig values
vm_data$total_revenue <- log(as.numeric(gsub("\\,", '', sub('\\$','',as.character(vm_data$`Total Revenue`)))))

vm_data$sentiment_clean <- round(as.numeric(vm_data$SentimentScore),5)

vm_data$InferredGender <- as.factor(vm_data$InferredGender)
vm_data$EmailTLD <- as.factor(vm_data$EmailTLD)
vm_data$Photo <- as.factor(vm_data$Photo)
vm_data$EmailDomain <- as.factor(vm_data$EmailDomain)
#vm_data$give_method_vlookup <- as.factor(vm_data$give_method_vlookup)
vm_data$give_method <- as.factor(vm_data$give_method)

# Drop old features

drop_cols <- c("SentimentScore"
               ,"Total Revenue"
               ,"votes_gallery_(Calc)"
               ,"votes_local_network")

vm_data <- vm_data %>% select(-one_of(drop_cols))

# Remove NA's in respose
vm_data <- vm_data %>% filter(!is.na(target))

vm_data2 <- na.roughfix(vm_data)

# Basic GBM for importance

vm_gbm2 <- gbm(target ~.
              ,data = vm_data2
              ,distribution = "poisson"
              ,n.trees = 1000
              ,interaction.depth = 5
              ,n.minobsinnode = 15
              ,shrinkage = 0.001
              ,bag.fraction = 0.6
              ,train.fraction = 0.5
              ,verbose = TRUE)

summary(vm_gbm2)

gbm.perf(vm_gbm2)


```


```{r email domain freq2}
vm_data2 %>% group_by(EmailDomain) %>%
  summarise(freq = n()) %>%
  filter(freq >= 25) %>%
  ggplot(aes(EmailDomain, freq)) + 
  geom_bar(stat="identity")

vm_data2 %>% group_by(EmailDomain) %>%
  summarise(avg_target = mean(target)
            ,freq = n()) %>%
  filter(freq >= 25) %>%
  ggplot(aes(EmailDomain, avg_target)) + 
  geom_bar(stat="identity")

```

Interestingly the GBM picks up on email domain which exhibits an interesting pattern with gallery vote count.

## Modeling network votes directly
```{r total vote count model}
set.seed(101)
# Extract columns to keep
keep_cols <- c("votes_local_network"
               ,"InferredGender"
               ,"TextLength"
               ,"NumWords"
               ,"NumPunctuation"
               ,"NumURLs"
               ,"NumHashtags"
               ,"NumAtMentions"
               ,"NumDollarsigns"
               ,"NumDigits"
               ,"NumExclamationpoints"
               ,"NumUppercaseletters"
               ,"EmailTLD"
               ,"ScrabbleScore"
               ,"Total Revenue"
               ,"Photo"
               ,"give_reason2"
               #,"give_method_vlookup"
               ,"give_method"
               ,"SentimentScore"
               ,"EmailDomain"
               )

vm_data <- gr_data %>%
  select(one_of(keep_cols))

# Target
vm_data$target <- vm_data$votes_local_network

# Some Cleaning
# Use total revenue on log scale due to some biiiig values
vm_data$total_revenue <- log(as.numeric(gsub("\\,", '', sub('\\$','',as.character(vm_data$`Total Revenue`)))))

vm_data$sentiment_clean <- round(as.numeric(vm_data$SentimentScore),5)

vm_data$InferredGender <- as.factor(vm_data$InferredGender)
vm_data$EmailTLD <- as.factor(vm_data$EmailTLD)
vm_data$Photo <- as.factor(vm_data$Photo)
vm_data$EmailDomain <- as.factor(vm_data$EmailDomain)
#vm_data$give_method_vlookup <- as.factor(vm_data$give_method_vlookup)
vm_data$give_method <- as.factor(vm_data$give_method)

# Drop old features

drop_cols <- c("SentimentScore"
               ,"Total Revenue"
               ,"votes_local_network")

vm_data <- vm_data %>% select(-one_of(drop_cols))

# Remove NA's in respose
vm_data <- vm_data %>% filter(!is.na(target))

vm_data2 <- na.roughfix(vm_data)

# Basic GBM for importance

vm_gbm2 <- gbm(target ~.
              ,data = vm_data2
              ,distribution = "poisson"
              ,n.trees = 1000
              ,interaction.depth = 5
              ,n.minobsinnode = 15
              ,shrinkage = 0.001
              ,bag.fraction = 0.6
              ,train.fraction = 0.5
              ,verbose = TRUE)

summary(vm_gbm2)

gbm.perf(vm_gbm2)


```

```{r email domain freq3}
vm_data2 %>% group_by(EmailDomain) %>%
  summarise(freq = n()) %>%
  filter(freq >= 25) %>%
  ggplot(aes(EmailDomain, freq)) + 
  geom_bar(stat="identity")

vm_data2 %>% group_by(EmailDomain) %>%
  summarise(avg_target = mean(target)
            ,freq = n()) %>%
  filter(freq >= 5) %>%
  ggplot(aes(EmailDomain, avg_target)) + 
  geom_bar(stat="identity")

```

## Modeling Total votes directly
```{r local network vote count model}
set.seed(101)
# Extract columns to keep
keep_cols <- c("votes_total"
               ,"InferredGender"
               ,"TextLength"
               ,"NumWords"
               ,"NumPunctuation"
               ,"NumURLs"
               ,"NumHashtags"
               ,"NumAtMentions"
               ,"NumDollarsigns"
               ,"NumDigits"
               ,"NumExclamationpoints"
               ,"NumUppercaseletters"
               ,"EmailTLD"
               ,"ScrabbleScore"
               ,"Total Revenue"
               ,"Photo"
               ,"give_reason2"
               #,"give_method_vlookup"
               ,"give_method"
               ,"SentimentScore"
               ,"EmailDomain"
               )

vm_data <- gr_data %>%
  select(one_of(keep_cols))

# Target
vm_data$target <- vm_data$votes_total

# Some Cleaning
# Use total revenue on log scale due to some biiiig values
vm_data$total_revenue <- log(as.numeric(gsub("\\,", '', sub('\\$','',as.character(vm_data$`Total Revenue`)))))

vm_data$sentiment_clean <- round(as.numeric(vm_data$SentimentScore),5)

vm_data$InferredGender <- as.factor(vm_data$InferredGender)
vm_data$EmailTLD <- as.factor(vm_data$EmailTLD)
vm_data$Photo <- as.factor(vm_data$Photo)
vm_data$EmailDomain <- as.factor(vm_data$EmailDomain)
#vm_data$give_method_vlookup <- as.factor(vm_data$give_method_vlookup)
vm_data$give_method <- as.factor(vm_data$give_method)

# Drop old features

drop_cols <- c("SentimentScore"
               ,"Total Revenue"
               ,"votes_total")

vm_data <- vm_data %>% select(-one_of(drop_cols))

# Remove NA's in respose
vm_data <- vm_data %>% filter(!is.na(target))

vm_data2 <- na.roughfix(vm_data)

# Basic GBM for importance

vm_gbm2 <- gbm(target ~.
              ,data = vm_data2
              ,distribution = "poisson"
              ,n.trees = 1000
              ,interaction.depth = 5
              ,n.minobsinnode = 15
              ,shrinkage = 0.001
              ,bag.fraction = 0.6
              ,train.fraction = 0.5
              ,verbose = TRUE)

summary(vm_gbm2)

gbm.perf(vm_gbm2)


```

```{r email domain freq4}
vm_data2 %>% group_by(EmailDomain) %>%
  summarise(freq = n()) %>%
  filter(freq >= 25) %>%
  ggplot(aes(EmailDomain, freq)) + 
  geom_bar(stat="identity")

vm_data2 %>% group_by(EmailDomain) %>%
  summarise(avg_target = mean(target)
            ,freq = n()) %>%
  filter(freq >= 5) %>%
  ggplot(aes(EmailDomain, avg_target)) + 
  geom_bar(stat="identity")

```